# Essential Biodiversity Variables

## SDM 
https://rspatial.org/raster/sdm/index.html

## Diversity 

Biological diversity analyses typically use multivariate techniques to assess variation in data sets comprising sampling events and cases. A sampling event can be across time and space. In biodiversity analysis three cases are the most common:

- Taxonomic diversity (species);
- Functional diversity (biological form); and

- Genetic diversity (allellic frequency, phylogeny etc).

Variation in these three dimensions can be directly compared both within and between these different dimensions of biological diversity. 

### Types of analyses - Patitioning $\alpha$, $\beta$, $\gamma$ *etc.* diversity

Two forms of diversity analyses are currently widely used; classic diversity measures (eg species richness, shannons diversity index *etc.*) and *numbers equivalents* representation of the underlying diversity distribution. This second form was first introduced in  Hill [-@hill_diversity_1973] and a number of resources are now available for computing both forms. See the forum piece *Forum: Partitioning diversity* in Ecology [-@ellison_partitioning_2010]  for a fuller discussion on the use of *numbers equivalents*. 

Both types of diversity partitioning are used for all types of analyses and we present resources which are available for both forms: 
 - Resources for classic diversity measures
 - Resources for *numbers equivalents*.

### Data form
Within the R environment both methods require data to be in the **wide format**. See [link to page explaining shift from long to wide].

```{r echo=FALSE}
library(kableExtra)

df <- data.frame(class_1 = sample(1:100,10,replace = TRUE),
                 class_2 = sample(1:100,10,replace = TRUE),
                 class_3 = sample(1:100,10,replace = TRUE),
                 class_4 = sample(1:100,10,replace = TRUE),
                 class_5 = sample(1:100,10,replace = TRUE),
                 class_6 = sample(1:100,10,replace = TRUE),
                 class_7 = sample(1:100,10,replace = TRUE),
                 class_8 = sample(1:100,10,replace = TRUE),
                 class_9 = sample(1:100,10,replace = TRUE),
                 class_10 = sample(1:100,10,replace = TRUE)
                 )
row.names(df) <- paste("Site",
                       1:nrow(df),sep="_")
kable(df, booktabs = TRUE, caption = "Example of wide format")%>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F)

```

### Classic diversity measures 

People have used many different indices to measure diversity. These include:

### $\alpha$ diversity

$\alpha$ diversity refers to the diversity at a single site. There are a number of different indices to caclculate the most common are:

- Species richness;
- Shannon/shannon weaver index;
- Simpson;
- Inverse Simpson; and
- Gini Simpson

The R libraries **vegan, adiv, abdiv** all provide methods to calculate these measures as well as a wealth of others. Within **abdiv** the funtion *alpha_diversity* lists the $\alpha$ diversity measures available within the package. Whilst not exhaustive it is a large list.

There are also ways of estimating $\alpha$ diversity through rarefaction as well as modelling and visualising its different aspects in both **vegan** and **adiv**.

### $\beta$ diversity and dissimilarities

The $\beta$ diversity is a measure of the change in composition and/or abundance between sites. There is a long history of methods to measure this particular aspect of diversity. This has resulted in multiple indices and dissimilarities. Commonly used indices include:

- Jaccard;
- Sørrensen;
- Bray-Curtis;
- Hellinger distance,
- Chord distance


An extensive list of $\beta$ diversities are available through the function *betadiver* in the **vegan** package as well as *beta_diversity* in the **abdiv** package. 

Methods for analysing $\beta$ diversity are included in the libraries **betapart, vegan, adiv and ade4**.

### Additional related diversities

Measures of dark diversity (those species that are absent from an ecosystem but which belong to its species pool, [@partel_dark_2011]) and $\zeta$ diversity (a measure of species pool overlap, [@hui_zeta_2014]) have been developed. These measures are instituted in the libraries:

- **DarkDiv** dark diversity; and
- **zetadiv** zeta diversity
 

### Numbers equivalents

Numbers equivalents account for the nested heirarchies in $\alpha$, $\beta$ and $\gamma$ diversity. A number of different libraries have now been developed:

- Hill numbers: @hill_diversity_1973
  - **adiv**
  - **HillR** 
  - **iNext**

- Tsallis entropy: @marcon_decomposing_2015
  - **adiv**
  - **entropart**
  
 - Crossing point theory: @patil_diversity_1982
    - **BioFTF**
    

### Example of analysis of diversity

In this example we use the sampling event "Vegetation data from sheep grazing experiment at alpine site in Hol, Norway" available [here](https://www.gbif.org/dataset/57d403f4-9d00-4651-8e26-1f4e81a21181). 

Downloading the Darwin core archive which contains a ... TO BE DONE...

```{r  echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
library(finch)
library(reshape2)
library(dplyr)

fileDirs <- list.dirs("./data/BDcleaner_Scripts/Example/Downloads/ignored data/Grazing experiments/Hol")
vegData <-  finch::dwca_read( list.files( fileDirs,
                     full.names = TRUE), read = TRUE)
```


#### Event data

```{r , echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
library(kableExtra)

kable(head(vegData$data$event.txt)) %>%
  kable_styling()
```

#### Occurence data

```{r , echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
library(kableExtra)

kable(head(vegData$data$occurrence.txt)) %>%
  kable_styling()
```
