# Query data from NILS: blueberry and mountain violet occurrence in alpine region over time

```{r setup_nils2, message=FALSE, warning=FALSE}

# Set up
library(jsonlite)
library(dplyr)
library(tidyr)
library(conflicted)
conflicts_prefer( dplyr::select, dplyr::filter)
library(glue)
library(ggplot2)
library(sf)
```

Following with NILS data we will now illustrate another example, this time from the report Skog & Mark 2015 [@naturvårdsverket2015]. In this example we reproduce the analys done in the first chapter with the aim to compare presence of two species at different altitudes in the mountain range.

## Query data

In this case we query the endpoint 'SmaprovytaArter' that contains the occurrence of different species per sample are.

```{r querypy, message=FALSE, warning=FALSE}
urlApi <- 'https://landskap.slu.se/api/nils/api/'
endpoint <- 'SmaprovytaArter'
years <- paste0('Ar=',c(2003:2012), collapse = '&')
region <- 'BioGeoRegion=Alpin'

pyUrl <- url(glue('{urlApi}{endpoint}?{years}&{region}'))
pyGet <- fromJSON(pyUrl)
py <- pyGet$data |> 
  mutate(uuid = glue('{rutaNummer}-{provytaNummer}')) |> 
  mutate_at(vars(uuid, vetenskapligtNamn), list(factor))

pySpp <- py |> 
  group_by(uuid, ar) |> 
  reframe(coordNS = unique(koordNS),
          coordEW = unique(koordEW),
          presBB = ifelse("Vaccinium myrtillus" %in% vetenskapligtNamn, 1, 0),
          presAV = ifelse("Viola biflora" %in% vetenskapligtNamn, 1, 0)) |> 
  as.data.frame()
```

In the same way we query NILS, there are many other sources of data. We query [opentopodata.org](opentopodata.org) the obtain the altitude at specific coordinates. But first we need to convert NILS sample areas coordinates into a coordinate system that this new service can understand (WGS84). And we want to make the query only one per sample area, as they wont go anywhere between sampling events.

```{r locations, message=FALSE, warning=FALSE}
pySf <- py |> 
  group_by(uuid) |> 
  reframe(coordNS = unique(koordNS),
          coordEW = unique(koordEW)) |>  
  rowwise() |> 
  mutate(geom = st_sfc(list(st_point(c(coordEW, coordNS))),
                       crs = st_crs(3006))) |> # SWEREF99 TM
  st_as_sf() |> 
  st_transform(st_crs(4326)) # WGS84

pySf$elevation <- NA

locations <- st_coordinates(pySf) |> 
  as.data.frame() |> 
  rename("latitude" = Y, 
         "longitude" = X)
```

We also want to iterate the query as this service has a quota of max 100 locations per call.

```{r altitude, eval=FALSE}
nLocations <- nrow(locations)
queryLimit <- 100
lChunks <- split(seq(nLocations), ceiling(seq(nLocations)/queryLimit))

urlAlt <- "https://api.opentopodata.org/v1/"
endpoint <- "eudem25m"

for (l in seq(length(lChunks))) {
  loc <- locations[lChunks[[l]],] |> 
    mutate("xy" = glue("{round(latitude,5)},{round(longitude,5)}"))
  loc <- paste0(loc$xy, collapse = "|")  
  elevGet <- url(glue("https://api.opentopodata.org/v1/eudem25m?locations={loc}"))
  elevation <- fromJSON(elevGet)
  # message(elevation$status)
  elevation <- elevation$results
  # print(range(elevation$elevation))
  
  pySf$elevation[lChunks[[l]]] <- elevation$elevation
  # print(l)
}
```

```{r savealt, eval=FALSE, echo=FALSE}
saveRDS(pySf, file = "data/07_nils2_pySf_with_elevation.rds")
```

```{r readalt, eval=TRUE, echo=FALSE}
pySf <- readRDS(file = "data/07_nils2_pySf_with_elevation.rds")
```

We put everything together (join) and summarise the proportion of sample areas per inventory phase and elevation group.

```{r summarisePY, message=FALSE, warning=FALSE, echo=FALSE}
pySf$elevationBin <- cut(pySf$elevation, 
                         breaks = c(seq(300,1000, 50),Inf),
                         include.lowest = TRUE, right = FALSE)

pySpp$invPha <- ifelse(as.numeric(as.character(pySpp$ar)) > 2007, 
                        "2008-2012", "2003-2007")

pySum <- pySpp |> 
  left_join(st_drop_geometry(pySf)) |> 
  group_by(invPha, elevationBin) |> 
  reframe(count = n(),
          sumPresBB = sum(presBB),
          sumPresAV = sum(presAV)) |> 
  mutate(occBB = sumPresBB/count*100,
         occAV = sumPresAV/count*100)

```

and finally, we create a beautiful plot.

```{r plotOcc, message=FALSE, warning=FALSE}
elevBinLabel <- c("300-349","350-399","400-449","450-499","500-549","550-599",
                  "600-649","650-699","700-749","750-799","800-849","850-899",
                  "900-949","950-999",">=1000")
#### Vaccinium myrtillus
ggplot(pySum, aes(x = elevationBin, y = occBB, color = invPha)) +
  geom_point(position = position_dodge(0.1)) +
  labs(title = "Occurrence of blueberry (Vaccinium myrtillus) \nin the mountains",
       y = "Occrrence (%)", x = "Altitude (m.a.s.l.)", 
       color = "Inventory phase") +
  ylim(0, 100) +
  scale_x_discrete(labels = elevBinLabel) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))

#### Viola biflora
ggplot(pySum, aes(x = elevationBin, y = occAV, color = invPha)) +
  geom_point(position = position_dodge(0.1)) +
  labs(title = 'Occurrence of alpine yellow-violet (Viola biflora) \nin the mountains', 
       y = "Occrrence (%)", x = "Altitude (m.a.s.l.)", 
       color = "Inventory phase") +
  ylim(0, 100) +
  scale_x_discrete(labels = elevBinLabel) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```
