---
title: "Data Completeness"
author: "Charlie Campbell"
date: "18/11/2020"
output: html_document
---
# Data completeness 

Knowledeg of the quality of data available is very important. Quanitfying gaps in data taxanomic, temporal and spatial is an important step. The R package **BIRDS** provides a resource to do all of these things. It can:

  - Summarise spatial distribution of records;
  - Summarise the temporal distribution of records; and
  - Thorugh the implementation of ignorance scores, summarise likely gaps in biodiversity knowledge.
  
  
```{r message=FALSE, warning=FALSE}
library(BIRDS)
data("bryophytaObs")
data("gotaland")

OB <- organiseBirds(bryophytaObs)

grid <- makeGrid(gotaland, gridSize = 10)

SB <- summariseBirds(OB, grid)

```

```{r message=FALSE, warning=FALSE, echo=FALSE}
library(sf)
library(cowplot)
library(ggplot2)
library(colorRamps)

spatial_sf <- st_as_sf(SB$spatial)

obs <- ggplot(data = spatial_sf ,aes( fill = nObs))+
  geom_sf()+
  ggtitle("Observations")+
  scale_fill_gradientn(colors = matlab.like2(100))+
  theme_cowplot()

Sp <- ggplot(data = spatial_sf ,aes( fill = nSpp))+
  geom_sf()+
  ggtitle("Number of species")+
  scale_fill_gradientn(colors = matlab.like2(100))+
  theme_cowplot()


obs
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
Sp
```