---
title: "Spatial data"
author: "Charlie Campbell"
date: "18/11/2020"
output: html_document
---
# Spatial data

Spatial data have coordinates. IN the simplest form they are a set of points. They may be polgonal (both regular and irregular) depicting any sort of feature. They may have additional information attached to them such as a data frame. They may take the form of a grid and may be in the raster format. Very frequently they represent the land surface and as such may have coordinate reference infomation attached (eg. SWEREF 99 TM) There are several resources for handling such data:

 - in R the most frequent base libraries encountered are:
 
    - **sp** - Classes and methods for spatial data 
    - **raster** - Geographic data analysis and modelling 
    - **sf** - Simple features for R 
    - **rgdal** -  Bindings for the 'Geospatial' Data Abstraction Library 
  
 - There are a number of other programs that are used for spatial GIS data. THese include:
    - QGIS - Open source GIS 
    - ArcGIS - Esri ArcGIS

# Exporting Data spatial data from R

Provided are the libraries and the appropriate functions for exporting spatial data from R

## Spatial vector data
```{r eval=FALSE}
library(sf)
write_sf() #capable of writing multiple different file formats

library(rgdal)
writeOGR() #capable of writing multiple different file formats

library(raster)
shapefile() #writes ESRI shapefiles
```
## Raster
```{r eval=FALSE}
library(raster)
writeRaster() #capable of writing multiple different file formats

```